\begin{section}{Parametricity}
\begin{frame}
\frametitle{Parametricity}
\begin{block}{Philip Wadler \cite{Wadler89theoremsfor} writes:}
\begin{quotation}
Write down the definition of a polymorphic function on a piece of paper. Tell me its type, but be careful not to let me see the function's definition. I will tell you a theorem that the function satisfies.

The purpose of this paper is to explain the trick.
\end{quotation}
\end{block}
\end{frame}

\begin{frame}[fragile]
\frametitle{A Simple Parametrically Polymorphic Function}
  Can you guess what this function does just by looking at it?
  \begin{minted}{scala}
    def doSomething[A](a: A, b: A): A = {
      //...
    }
  \end{minted}
\end{frame}

\begin{frame}{How did we do it?}
  \begin{center}
    \begin{itemize}
    \item Power of parametric polymorphism
    \item Very little power for the implementor
    \item Lot of info available to the reader from definition
    \end{itemize}
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Another (familiar?) Parametrically Polymorphic Function}
  Can you guess what this function does just by looking at it?
  \begin{minted}{scala}
    def doSomethingElse[A, B](a: List[A])(f: A => B): List[B] =
    {
      //...
    }
  \end{minted}
\end{frame}

\begin{frame}[fragile]{But wait...}
  \begin{minted}{scala}
    def doSomethingElse[A, B](a: List[A])(f: A => B): List[B] =
    {
      List(f(a.head))
    }
  \end{minted}
\end{frame}

\begin{frame}{What Happened There?}
  \begin{itemize}
    \item Parametric polymorphism alone isn't enough
    \item Implementor sometimes still has too much info to work with
    \item Reader sometimes still has to look under the covers to feel really safe
  \end{itemize}
\end{frame}

\end{section}
